Bayesian exponential random graph modeling of KWL Burial network

### 2020-12-09 analysis/paper/paper.Rmd
udpate discussion section

<!-- we should tell the reader what does it mean high-medium-low --> 
We analyzed burial data collected from the published excavation reports, and the original fieldwork notes of Kiwulan. Gold-foil and agate beads are commonly found across burials with substantial differences in quantities, so we described each burial as having one of three levels, high (> 10), medium (> 1 and < 10), and low (= 1), of these beads. Glass beads are in levels of high (> 100), medium (> 99 and < 900), and low (< 100). That means if burial 1 and burial 2 both have high quantities of agate beads, then there will be a tie connecting them. For less frequent prestige goods, including imported porcelains, gold foils, and fish-shaped ornaments, we linked two burials when they possess each type of goods in common (i.e. presence or absence). Node attributes here include osteological data, such as age and sex, and cultural data, such as ritual pottery, and burial value index. Ritual pottery was identified as locally made ceramics placed above graves that suggests funeral feasting according to historical records. The burial value index is an important attribute for economical inequality that we assigned the burials into three classes, high (>30), medium (>11 and <30), and low (<12), as an index of wealth. The number is the sum of values of each type of prestige good, which is calculate by the total number of a prestige item from burial context over the number of prestige item in a burial. Since burials tend to have multiple prestige goods in common, the network ties are weighted instead of binary data (where the value 1 represents a tie and the value 0 otherwise) [@Snijders2011]. For example, if two burials have both low quantity of glass beads and porcelain in common, the tie is given a value of 2. Our networks are non-directed, which means ties have no orientation forming the relationship between actors to indicate a mutual relationship.

@@

<!-- theoretical framework for the results-->
A striking finding in our results is the change in network properties of Kiwulan burials from a more cohesive network (represented by higher transitivity and negative popularity) toward a more centralized network (lower transitivity and positive popularity) after the European presence during the 17th century. This supports our hypothesized models of chronological change from a more corporate society to a more networked society, as indicated in the grave goods and burial attributes. These observations are in line with theoretical assumptions of horizontal hierarchy, the corporate-network continuum that focuses on distinct strategies for achieving power as an alternative explanation to understand the emergence and level of social inequality [@Drennan2010; @Feinman2000]. Compared to traditional models of vertical hierarchies that stress stratified social characteristics, the corporate-network model may be more relevant for understanding how a small scale society was organized, such as we see here in the pericolonial context of Kiwulan. A corporate-based society, such the pre-European situation at Kiwulan, stresses shared power between individuals, integrative rituals, and social inequality, if any, would be associated with a social group or sector [@Siegel1999]. In contrast, a more network-based society, indicated by the post-European data from Kiwulan, presents wealth accumulation through individual networks, resource control, and trade monopolization.

@@

<!-- other evidence and a bigger anthropological picture-->
We used burial data as a proxy to explore the social relations based on the argument that burials could be viewed as a social representation to understand social structures [@Binford1971; @Saxe1970]. We note that there could be an issue of disconnection between the role of the deceased in a burial context and their real life, which should be reduced by examining evidence from the residential context [@Chapman2003]. Previous studies of trade ornaments from the residential area of Kiwulan suggests an uneven spatial distribution since the Spanish and the Dutch were active in northern Taiwan that hints at an increased social differentiation [@LWandBM2020ornament]. This is consistent with our result of a more centralized burial network after the presence of the Europeans, which supports the assumption of social role representation from burials in this case. Since the burial goods used for making ties were treated as prestige goods in local Indigenous culture [@Borao2009; @LiandWu2006], the interrelations indicated by the flow of goods could reflect the process of social changes at Kiwulan in a pericolonial context. After the presence of the Europeans, a more centralized and less cohesive network indicates some individuals were differentiated or privileged to possess more distinctive goods. Moreover, the relationship tends to form between individuals with the same wealth level represented by the rarity of goods. This uneven prestige accumulation might result from differential access to exotic trade relative to the European colonization. The involvement in long-distance exchange often leads to changes in political-economic strategies toward a network mode, especially for small-scale societies, where ambitious individuals are able to build their factions and personal influence through the distribution of high value goods [@Feinman2000].

@@

# Conclusion

>>>>

<!-- burial data and the definition of tie formation  --> 
We analyzed burial data collected from the published excavation reports, and the original fieldwork notes of Kiwulan. Trade beads are commonly found across burials with substantial differences in quantities, so we described each burial as having one of three levels, high, medium, and low, according to their distributions across all burials. Gold-foil beads are in levels of high (>3), medium (=2-3), and low (=1); glass beads and carnelian beads are in levels of high (>6), medium (=3-6), and low (=1-2). That means if burial 1 and burial 2 both have high quantities of agate beads, then there will be a tie connecting them. For less frequent prestige goods, including imported porcelains, gold foils, and fish-shaped ornaments, we linked two burials when they possess each type of goods in common (i.e. presence or absence). Node attributes here include osteological data, such as age and sex, and cultural data, such as ritual pottery, and burial value index. Ritual pottery was identified as locally made ceramics placed above graves that suggests funeral feasting according to historical records. The burial value index is an important attribute for economical inequality that we assigned the burials into three classes, high (>30), medium (>11 and <30), and low (<12), as an index of wealth. The number is the sum of values of each type of prestige good, which is calculate by the total number of a prestige item from burial context over the number of prestige item in a burial. Since burials tend to have multiple prestige goods in common, the network ties are weighted instead of binary data (where the value 1 represents a tie and the value 0 otherwise) [@Snijders2011]. For example, if two burials have both low quantity of glass beads and porcelain in common, the tie is given a value of 2. Our networks are non-directed, which means ties have no orientation forming the relationship between actors to indicate a mutual relationship.

@@

<!-- theoretical framework for the results-->
A striking finding in our results is the change in network properties of Kiwulan burials from a  more cohesive network with multiple subgroups (represented by higher transitivity and negative popularity) toward a more centralized network with  (lower transitivity and positive popularity) after the European presence during the 17th century. This supports our hypothesized models of chronological change from a more corporate society to a more networked society, as indicated in the grave goods and burial attributes. These observations are in line with theoretical assumptions of horizontal hierarchy, the corporate-network continuum that focuses on distinct strategies for achieving power as an alternative explanation to understand the emergence and level of social inequality [@Drennan2010; @Feinman2000; @Feinman2000political]. Compared to traditional models of vertical hierarchies that stress stratified social characteristics, the corporate-network model may be more relevant for understanding how a small scale society was organized, such as we see here in the pericolonial context of Kiwulan. A corporate-based society, such as the pre-European situation at Kiwulan, stresses shared power between individuals, communal  rituals, and social inequality, if any, would be associated with a group association [@Siegel1999]. In contrast, a more network-based society, indicated by the post-European data from Kiwulan, presents wealth accumulation through individual networks, resource control, and trade monopolization.

@@

<!-- other evidence and a bigger anthropological picture-->
We used burial data as a proxy to explore the social relations based on the argument that burials could be viewed as a social representation to understand social structures [@Binford1971; @Saxe1970]. We note that there could be an issue of disconnection between the role of the deceased in a burial context and their real life, which should be reduced by examining evidence from the residential context [@Chapman2003]. Previous studies of trade ornaments from the residential area of Kiwulan suggests an uneven spatial distribution when the Europeans were active in northern Taiwan that hints at an increased social differentiation at Kiwulan in a pericolonial context [@LWandBM2020ornament]. This is consistent with the result of a more centralized burial network after the presence of the Europeans, which supports the assumption of social role representation from burials in this case. Since the burial goods used for making ties were treated as prestige goods in local Indigenous culture [@Borao2009; @LiandWu2006], the interrelations represented by the flow of goods could reflect the process of social changes at Kiwulan in a pericolonial context. After the arrival of Europeans, some individuals may be differentiated or privileged to possess more distinctive goods through becoming the trade partners with the Europeans. According to historical records, the Dutch would give local Indigenous leaders symbolic items or trade goods to assert their authority and consolidate the relationship with local people [@Kang2016, pp137-162]. Uneven distribution of goods in burials we observed could result from differential access to exotic trade relative to the European colonization and gradually led to increased social differentiation. The involvement in long-distance exchange often leads to changes in political-economic strategies to a network mode, where ambitious individuals are able to build their factions and personal influence through the distribution of high value goods [@Feinman2000; ].

@@

# Conclusion
It should be noted that corporate/network spectrum represents a different degree of hierarchical complexity instead of a clear division. 
This might lead to accumulation of rare prestige goods and competition between individuals that Kiwulan gradually moves to a more network-based society.
